Carbon stocks in aboveground biomass for Colombian mangroves with associated uncertainties

### 2017-06-05 Article_mangroves.tex

\section{Methods}
\subsection{Data}
We used carbon data from two different sources: (i) scientific articles and/or technical reports from studies developed in mangrove areas located in the Caribbean and Pacific coast of Colombia, in which above ground biomass (AGB) was quantified and; (ii) data from two inventories of carbon stocks in Cispat\'{a} bay (Colombian Caribbean coast) and M\'{a}laga bay (Colombian Pacific coast) \citep{Bolivar_inpreparation, Malaga2015}. For the first type of data, we selected only studies which included coordinates or specific location names that can be easily geolocated \citep{INVEMAR2007, Lema2007, DelaPena2010, Blanco2012}. The main difference between both data types is the scale of the information. The first data type report average AGB per location, while the second data type reports estimates per plot covering the same region  (22 permanent plots of 500 m\textsuperscript2 in Cispat\'{a} bay  and 10 in M\'{a}laga bay).

Bioclimatic data were extracted from the WorldClim Bioclim \citep{Hijmans2005} 30 arc-second data set (http://www.worldclim.org/bioclim). Vegetation condition data consisted of the Enhanced Vegetation Index values extracted from the MOD13Q1 MODIS Vegetation Indices Product. The product represents 16-day composites at 232 m spatial resolution. We selected EVI over the Normalized Difference Vegetation Index, also included in the MOD13Q1 product because it offers better sensitivity to forest canopy variations, specially in areas with high biomass (Huete et al 2002. \marginpar{Huete, A., K. Didan, T. Miura, E. P. Rodriguez, X. Gao, and L. G. Ferreira. 2002. Overview of the radiometric and biophysical performance of the MODIS vegetation indices. Remote Sensing of Environment 83:195-213.}
\end{comment}) over Likewise, global MODIS vegetation indexes provide consistent spatial and temporal comparisons of vegetation conditions; in this sense, EVI values were retrieved from the online data pool, courtesy of the NASA Land Processes Distributed Active Archive Center (LP DAAC) \citep{MODIS2015} (https://lpdaac.usgs.gov/data\_access/data\_pool). 
AGB data sources used to develop the predictive models are provided in the supplementary material.

>>>>

\section{Methods}
\subsection{Data}
We used carbon data from two different sources: (i) scientific articles and/or technical reports from studies developed in mangrove areas located in the Caribbean and Pacific coast of Colombia, in which above ground biomass (AGB) was quantified and; (ii) data from two inventories of carbon stocks in Cispat\'{a} bay (Colombian Caribbean coast) and M\'{a}laga bay (Colombian Pacific coast) \citep{Bolivar_inpreparation, Malaga2015}. For the first type of data, we selected only studies which included coordinates or specific location names that can be easily geolocated \citep{INVEMAR2007, Lema2007, DelaPena2010, Blanco2012}. The main difference between both data types is the scale of the information. The first data type report average AGB per location, while the second data type reports estimates per plot covering the same region  (22 permanent plots of 500 m\textsuperscript2 in Cispat\'{a} bay  and 10 in M\'{a}laga bay). AGB data sources used to develop the predictive models are provided in the supplementary material.

Bioclimatic data were extracted from the WorldClim Bioclim \citep{Hijmans2005} 30 arc-second data set (http://www.worldclim.org/bioclim).

Vegetation condition data consisted of the Enhanced Vegetation Index values extracted from the MOD13Q1 MODIS Vegetation Indices Product. The product represents 16-day composites at 232 m spatial resolution. We selected EVI over the Normalized Difference Vegetation Index, also included in the MOD13Q1 product because it offers better sensitivity to forest canopy variations, specially in areas with high biomass (Huete et al 2002) \marginpar{Huete, A., K. Didan, T. Miura, E. P. Rodriguez, X. Gao, and L. G. Ferreira. 2002. Overview of the radiometric and biophysical performance of the MODIS vegetation indices. Remote Sensing of Environment 83:195-213.}. EVI values were retrieved from the online data pool, courtesy of the NASA Land Processes Distributed Active Archive Center (LP DAAC) \citep{MODIS2015} (https://lpdaac.usgs.gov/data\_access/data\_pool). The MOD13 product comes with three levels of pixel reliability. For our purpose, we extracted all pixel values with good and marginal reliability for the years 2011??? through 2013???. These years cover the time range in which most of the carbon plots considered for the analysis were established. We included the marginal reliability pixels because the number of good quality pixels is not engough to cover the whole study area, given they high cloud cover throgout the year, specially in the pacific coast. After pixels values were extracted, we excluded pixels that had a reflectance in the blue band higher than 10 percent because these pixels are tipically associated to clouds. We also removed unreliable pixels based on abrupt drop or spikes in the data time series (Gutierrez-Velez and DeFries 2013) \marginpar{Gutiérrez-Vélez, V. H., and R. DeFries. 2013. Annual multi-resolution detection of land cover conversion to oil palm in the Peruvian Amazon. Remote Sensing of Environment 129:154-167}. Finally, we derived  the mean EVI per pixel within the study area.
